---
title: "Prediction Assignment Writeup"
author: "helloausrine"
date: "March 3, 2018"
output: rmd_document
---

### Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

###  Getting the Data

```{r}
trainingUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(trainingUrl, destfile = "./data/training.csv", method = "curl")
download.file(testingUrl, destfile = "./data/testing.csv", method = "curl")
training <- read.csv("./data/training.csv", na.strings = c("NA", "#DIV/0!"))
testing <- read.csv("./data/testing.csv", na.strings = c("NA", "#DIV/0!"))
```

### Examining the Data

```{r}
library(dplyr)
View(training)
View(testing)
```

The training set has 19622 observations of 160 variables while in the testing set there are 20 observations of 160 variables. There are a lot of NAs present in the data. 

### Cleaning the Data

```{r}
goodCol <- colSums(is.na(training)) < 1900
myTraining <- training[ , goodCol][ , ]
myTesting <- testing[ , goodCol][ , ]
myTraining <- myTraining[ , -(1:7)]
myTesting <- myTesting[ , -(1:7)] 
View(myTraining)
View(myTesting)
```

Here variables that contained many NA values were removed in order to clean the data.

### Subsetting the Training Data

```{r}
library(caret)
set.seed(4848)
inTrain <- createDataPartition(y = myTraining$classe, p = 0.75, list = FALSE)
inTraining <- myTraining[inTrain, ]
inTesting <- myTraining[-inTrain, ]
```

### Building the Model

```{r}
library(randomForest)
set.seed(555)
rfGrid <-  expand.grid(interaction.depth = c(1, 5, 9),
                        n.trees = (1:30)*50,
                        shrinkage = 0.1)
modelFit <- randomForest(classe ~., data = inTraining, tuneGrid = rfGrid) 
print(modelFit)
plot(modelFit)
```

### Cross Validation
 
```{r}
predictions <- predict(modelFit, newdata = inTesting)
confusionMatrix(predictions, inTesting$classe)
```

The model is confirmed to be sound in terms of sensitivity and accuracy.

### Final Validation 

```{r}
answers <- predict(modelFit, newdata = myTesting, type = "response")
print(answers)
```

### Conclusion

As seen in the final validation, the 20 answers were indeed correct.
